Well-quasi-orderings on word languages

### 2025-10-07 src/infixes.tex
Remove erroneous statement in Thm 2
Simplify Lemma 3

Let us first remark that if $S$ is a \kl{chain} for the \kl{suffix relation}
and $P$ is a \kl{chain} for the \kl{prefix relation}, then $SP$ is
\kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
right-to-left implication of \cref{bounded-language:thm}. Furthermore, any
language $L$ included in a finite union of products $S_i \cdot P_i$
is in fact a \kl{bounded language}. \todo[inline]{I think we can in fact construct chains that are not bounded languages. E.g. $ab, aba^2b, aba^2ba^3b, \ldots$?}

\todo[inline]{aliaume: yes, this way nonsense.}

@@

\begin{lemma}
    \label{periodic-infixes:lem}
    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
    having \kl{periodic lengths} $p$ and $q$ respectively.
    Then, if $u \infleq v$ and $\card{u} \geq pq$,
    then $u$ and $v$ share the same \kl{periodic length}
    $p = q$.
\end{lemma}
\begin{proof}
    The fact that $u$ and $v$ are \kl{periodic length}
    respectively $p$ and $q$ translates into the fact that $u_{i+p} = u_i$ and
    $v_{i+q} = v_i$ for all indices $i \in \Nat$ such that those letters are
    well-defined.

    Now, assume that $u$ is an \kl{infix} of $v$, this provides the existence
    of a $k \in \Nat$ such that $u = v_{k} \cdots v_{k + \card{u} - 1}$. In
    particular, $v_{k+i+p} = v_{k+i}$ for all $i \in \Nat$ such that $k+i+p < k
    + \card{u}$. Due to the periodicity of $v$, we also have $v_{k+i+q} = v_{k+i}$ for all $i$ with $k + i + q \leq k + \card{u}$. We conclude that both $u$ and $v$ are of
    \kl{periodic length} the greatest common divisor of $p$ and $q$, and by
    minimality of $q$ this must be equal to $q$ and to $p$.
    \todo{maybe elaborate on why $pq$ as bound on $\card{u}$ (also technically $2\max\{p,q\}$ would be enough, but maybe more cumbersome?}
\end{proof}

>>>>

Let us first remark that if $S$ is a \kl{chain} for the \kl{suffix relation}
and $P$ is a \kl{chain} for the \kl{prefix relation}, then $SP$ is
\kl{well-quasi-ordered} for the \kl{infix relation}. This proves the (easy)
right-to-left implication of \cref{bounded-language:thm}.

@@

\begin{lemma}
    \label{periodic-infixes:lem}
    Let $u,v \in \Sigma^*$ be two (non-empty) \kl{periodic words}
    having \kl{periods} of length $p$ and $q$ respectively.
    If $u \infleq v$ and $\card{u} \geq 2\max\{p, q\}$,
    then $u$ and $v$ both have a period of length
    $\gcd(p,q)$.
\end{lemma}
\begin{proof}
    Since $u$ has a period of length $p$ we have $u_{i+p} = u_i$ for all indices $i \in \Nat$ such that those letters are
    well-defined. Since $u$ is an \kl{infix} of $v$ and $\card{u} \geq q$, we also have $u_{i+q} = u_i$. It follows that $u$ has a period of length $\gcd(p, q)$.

    Since $v$ contains $u$ as a suffix and $\card{u} \geq q$, we can also conclude that $v$ must also have period of length $\gcd(p,q)$.
\end{proof}
